---
title: "4.X_visualizeStandard"
format: 
  html:
    theme: flatly
    toc: true
    toc-depth: 4
    fig-format: svg 
    embed-resources: true
    self-contained-math: true
editor: source
execute: 
  echo: false
  warning: false
  error: false
---



Visualize standard methods' results (Neg Bin, Linear on transformed, and SCEPTRE). 
These standard methods should have false positives when not including covariates (the confounders) and should have none or few false positives when including the covariates.





# Load


```{r output=FALSE}
# Libraries and Parameters
# ========================================================================================================================================================================
suppressPackageStartupMessages(library(assertthat)) # for some assert statements
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(ggh4x))


theme_set(theme_cowplot() +
            theme(plot.title = element_text(hjust = .5),
                  plot.subtitle = element_text(hjust = .5),
                  strip.background = element_rect(fill = "white", color = 'black'),
                  panel.grid.minor = element_blank(),
                  panel.grid.major = element_blank()))


# === === === === === === === === === === === === === === === === === === === === === === === === === 
# parameters
DEVICE                = 'macbook'
AYZW_setting_name     = 'B'
NC_names_raw = 'simpleLinear'
NC_names = strsplit(x = NC_names_raw, split = '-') |> unlist()
proximal_setting_name_count = 'simpleCount'
source('../PATHS.R') # load in data_dir and save_dir and CODE_DIR, depending on DEVICE value

# --- Folder ---
plot_dir = sprintf('%s/AY/%s/plots/standard/', save_dir, AYZW_setting_name)
dir.create(plot_dir, showWarnings = FALSE, recursive=TRUE)

# === === === === === === === === === === === === === === === === === === === === === === === === === 

# load chosen AYZW names
AY   = read.csv(sprintf('%s/AY/%s/AY.csv', save_dir, AYZW_setting_name))

# Continuous
# ========================================================================================================================================================================
effects_continuous = read.csv(sprintf('%s/AY/%s/%s/effects_continuous.csv', save_dir, AYZW_setting_name, NC_names_raw))
# effects_continuous = effects_continuous |> filter(type != 'maybe')
# effects_continuous$type = factor(effects_continuous$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))

# Count GLM
# ========================================================================================================================================================================
effects_count = read.csv(sprintf('%s/AY/%s/%s/effects_count.csv', save_dir, AYZW_setting_name, proximal_setting_name_count)) 
# effects_count = effects_count |> filter(type != 'maybe')
# effects_count$type = factor(effects_count$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


# SCEPTRE
# ========================================================================================================================================================================
# read in the previously formatted relevant rows when creating the AY pairs (2.3_xxx.r)
effects_sceptre = read.csv(sprintf('%s/AY/%s/sceptre_AY.csv', save_dir, AYZW_setting_name)) 
# effects_sceptre = effects_sceptre |> filter(type != 'maybe')
# effects_sceptre$type = factor(effects_sceptre$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


# head(effects_sceptre); colnames(effects_sceptre)
# colnames(effects_count)
# format to other effects df format
# doesn't really matter, but for now, we set the
# ATE estimate = log_2_fold_change! and the se is NA for now.
# because we have se(fold_change) but maybe needs some more transofrmation for log_2
effects_sceptre = effects_sceptre |>
  dplyr::rename(pval = p_value, ATE = log_2_fold_change) |>
  dplyr::mutate(se = NA, time_sec = NA, NC_type = NA,
                numNC = NA, basis = NA
                )

effects_sceptre = effects_sceptre[, colnames(effects_sceptre) %in% colnames(effects_continuous)]








# Config order and colors
# ========================================================================================================================================================================
# for each method, color that represents. also use this ordering
method_color_df = data.frame(method=c('lmYA', 'lmYAU', 
                                        'poisYA', 'poisYAU', 
                                        'nbYA', 'nbYAU', 
                                        'sceptre', 'sceptreU'),
                              color =c(colorRampPalette(c('white', 'darkblue'))(8)[c(4, 8)],   
                                       colorRampPalette(c('white', 'darkgreen'))(8)[c(4, 8)],
                                       colorRampPalette(c('white', 'darkorchid4'))(8)[c(4, 8)],
                                       colorRampPalette(c('white', 'darkorange2'))(8)[c(4, 8)]
                             ))

# 2 lines for (w/ confounders)
method_nicenames_multiline = c('lmYA'    = 'Linear',
                     'lmYAU'   = 'Linear\n(w/ confounders)',
                     'poisYA'  = 'Poisson',
                     'poisYAU' = 'Poisson\n(w/ confounders)',
                     'nbYA'    = 'Negative Binomial',
                     'nbYAU'   = 'Negative Binomial\n(w/ confounders)',
                     'sceptre' = 'SCEPTRE',
                     'sceptreU' = 'SCEPTRE\n(w/ confounders)')
# 1 line
method_nicenames = c('lmYA'    = 'Linear', 
                     'lmYAU'   = 'Linear (w/ confounders)', 
                     'poisYA'  = 'Poisson', 
                     'poisYAU' = 'Poisson (w/ confounders)', 
                     'nbYA'    = 'Negative Binomial', 
                     'nbYAU'   = 'Negative Binomial (w/ confounders)', 
                     'sceptre' = 'SCEPTRE',
                     'sceptreU' = 'SCEPTRE (w/ confounders)')
method_color_df$method_nicename = unname(method_nicenames)

method_colors = method_color_df$color |> `names<-`(method_color_df$method_nicename)

# barplot(1:length(method_color_df$color), col = method_color_df$color)


# All
# ========================================================================================================================================================================
effects = dplyr::bind_rows(effects_continuous, effects_count, effects_sceptre)
# effects = effects |> filter(type != 'maybe')
effects$type = factor(effects$type,
       levels = c('negative', 'positive', 'maybe'),
       labels = c(    'Null', 'Alternative', 'Discovery'))

# effects = effects |> mutate(method_long = paste0(method, numNC))
# effects |> arrange(A, Y, method)

effects = effects |> 
          mutate(method_raw       = method, 
                 method           = factor(method_raw, levels = method_color_df$method, labels = method_color_df$method_nicename),
                 method_multiline = factor(method_raw, levels = names(method_nicenames_multiline), labels = method_nicenames_multiline)) 
effects = effects |> select(-NC_type, -numNC, -basis)
```




# Compare standard methods


# ATE Pairplots
```{r fig.width=10, fig.height=10}

effects  |> select(A, Y, method_raw, ATE) |> reshape2::dcast(A + Y ~ method_raw, value.var = 'ATE') |> select(-A, -Y) |> pairs(main = 'ATE')
effects  |> select(A, Y, method_raw, pval) |> reshape2::dcast(A + Y ~ method_raw, value.var = 'pval') |> select(-A, -Y) |> pairs(main = 'p-value')

# effects |> filter(method_raw %in% c('nbYAU', 'sceptreU')) |> select(A, Y, method_raw, ATE) |> reshape2::dcast(A + Y ~ method_raw, value.var = 'ATE') |> ggplot() + geom_abline(aes(intercept = 0 , slope = 1), col = 'darkgray') + geom_point(aes(x = nbYAU, y = sceptreU))  + labs(title = 'NB vs SCEPTRE (w/ U)')
# 
# effects |> filter(method_raw %in% c('nbYAU', 'lmYAU')) |> select(A, Y, method_raw, ATE) |> reshape2::dcast(A + Y ~ method_raw, value.var = 'ATE') |> ggplot() + geom_abline(aes(intercept = 0 , slope = 1), col = 'darkgray') + geom_point(aes(x = nbYAU, y = lmYAU))  + labs(title = 'NB vs Lin (w/ U)')
```



# Standard Method Results

## Histogram of estimates and p-values

Histogram of estimates
```{r, fig.width=10, fig.height = 10}

method_multiline_labeller <- function(x) {
  y = c()
  for(x_ in x) {
    if(x_ %in% names(method_nicenames_multiline)) {y = c(y, method_nicenames_multiline[x_])}
    else {y=c(y,x_)}
  }
  return(y)
}
p = ggplot(effects,
       aes(x = ATE, fill = method)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .05) + # bins = 50) +
  geom_vline(aes(xintercept = 0), color = 'gray15', linewidth = .4) +
  coord_cartesian(xlim = c(-2, 1), ylim = c(0, 3)) +
  scale_y_continuous(breaks = 0:2, expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_colors, 
                    guide = guide_legend(ncol = 1)) +
  facet_grid(cols = vars(type), 
             rows = vars(method_raw),  # vars(method_long), # vars(method, NC_type, numNC), 
             labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
             scales = 'free') + 
  labs(title = 'Histogram of Estimates',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))
p
ggsave(sprintf('%s/estimates_hist_tall.pdf', plot_dir), height = 10, width= 10)

```

```{r fig.width=12, fig.height = 5}
p + facet_grid(rows = vars(type), 
             cols = vars(method_raw),  # vars(method_long), # vars(method, NC_type, numNC), 
             labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
             scales = 'free')  +
  theme(strip.text.x = element_text(size = 8), 
        axis.text.x = element_text(size = 8), 
        panel.spacing.y = unit(1, "lines"),
        panel.spacing.x = unit(0, "lines"))
ggsave(sprintf('%s/estimates_hist_wide.pdf', plot_dir), height = 5, width= 12)
```


Histogram of p-values
```{r fig.width=10, fig.height = 10}

p = ggplot(effects,
       aes(x = pval, fill = method)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .020) + # bins = 50) +
  scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1'), 
                     expand = expansion(mult = c(.01, .01))) +
  # coord_cartesian(ylim = c(0, 3)) +
  scale_y_continuous( expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_colors, guide = guide_legend(ncol = 1)) +
  facet_grid2(cols = vars(type), 
             rows = vars(method_raw), # vars(method_long), # vars(method, NC_type, numNC), 
             labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
             axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
             scales = 'free') + 
  labs(title = 'Histogram of p-values',
       x = 'p-value',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))
p

ggsave(sprintf('%s/pval_hist_tall.pdf', plot_dir), height = 10, width= 10)


# # # force same axes- does not look good, but gives perspective
# p + facet_grid2(cols = vars(type),
#              rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC),
#              axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
#              scales = 'fixed')
```


```{r fig.width=15, fig.height = 5}
p + 
# facet_grid(rows = vars(type), 
#              cols = vars(method_raw),  # vars(method_long), # vars(method, NC_type, numNC), 
#              labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
#              scales = 'free')  +
    facet_grid2(rows = vars(type), 
                cols = vars(method_raw), # vars(method_long), # vars(method, NC_type, numNC), 
             labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
             axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
             scales = 'free') + 
  theme(strip.text.x = element_text(size = 8), 
        axis.text.x = element_text(size = 8), 
        panel.spacing.y = unit(1, "lines"),
        panel.spacing.x = unit(0, "lines"))
ggsave(sprintf('%s/pval_hist_wide.pdf', plot_dir), height = 5, width= 15)
```


```{r fig.width=15, fig.height=5}
# force same axes- does not look good, but gives perspective
p + 
  facet_grid(rows = vars(type),
             cols = vars(method_raw),  # vars(method_long), # vars(method, NC_type, numNC),
             labeller = labeller(method_raw=as_labeller(method_multiline_labeller)),
             scales = 'free')  +
  theme(strip.text.x = element_text(size = 8), 
        axis.text.x = element_text(size = 8), 
        panel.spacing.y = unit(1, "lines"),
        panel.spacing.x = unit(0, "lines"))
ggsave(sprintf('%s/pval_hist_wide2.pdf', plot_dir), height = 5, width= 12)
```


## QQ-plots/Rejection Rates of p-values


QQ-plot (basically)
```{r fig.height=3, fig.width=9}
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects |> # filter(method_raw %in% c('sceptre', 'sceptreU')) |> 
            group_by(type, method) |> 
            reframe(pval=sort(pval), count=sum(!is.na(pval)), rejection_rate = (1:count)/count) 

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .5, alpha = .9, key_glyph = "rect") +
  geom_point(size=.4, alpha=.7, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(cols = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir), height = 3, width= 9)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval+(1e-500), base = 10), x = -log(rejection_rate+(1e-500), base = 10), color = method)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .5, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  facet_grid(cols = vars(type), scales = 'free') + # facetted_pos_scales doesn't seem to work with facet_grid
  # facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log.pdf', plot_dir), height = 3, width= 9)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.2, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.2, 2)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(.2, 2))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 10), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 10), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 10), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.pdf', plot_dir), height = 3, width= 9)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + geom_line() + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 3)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 3)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 3))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail2.pdf', plot_dir), height = 3, width= 9)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'


ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir), height = 3, width= 9)
```




# SCEPTRE

## With and Without Confounders (U)

```{r fig.width=9, fig.height=3}
plot_df = effects |> filter(method_raw %in% c('sceptre', 'sceptreU')) |>
            group_by(type, method) |>
            reframe(pval=sort(pval), count=sum(!is.na(pval)), rejection_rate = (1:count)/count)


# commpare with and without confounders
ggplot(plot_df, aes(x = pval, y = rejection_rate, color = method)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  geom_point(size=.4, alpha=.7, key_glyph = "rect") +
  # geom_line(linewidth = 1, alpha = .9, key_glyph = "rect") +
  scale_color_manual(values = method_colors, guide = guide_legend(ncol = 1)) +
  facet_grid(cols = vars(type))


```


```{r fig.width=12, fig.height = 4}
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects |> filter(method_raw %in% c('sceptre', 'sceptreU')) |> 
            group_by(type, method) |> 
            reframe(pval=sort(pval), count=sum(!is.na(pval)), rejection_rate = (1:count)/count) 

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(size=.4, alpha=.7, key_glyph = "rect") +
  # geom_line(linewidth = 1, alpha = .9, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(cols = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir), height = 7, width= 8)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .9, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  facet_grid(cols = vars(type), scales = 'free') + # facetted_pos_scales doesn't seem to work with facet_grid
  # facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log.pdf', plot_dir), height = 7, width= 8)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.2, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.2, 2)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(.2, 2))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 40), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 40), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.pdf', plot_dir), height = 6, width= 6)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 3)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 3)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 3))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail2.pdf', plot_dir), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'


ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir), height = 6, width= 6)
```



















