---
title: "4.X_visualizeStandard"
format: 
  html:
    theme: flatly
    toc: true
    toc-depth: 4
    fig-format: svg 
    embed-resources: true
    self-contained-math: true
editor: source
execute: 
  echo: false
  warning: false
  error: false
---



Visualize standard methods' results (Neg Bin, Linear on transformed, and SCEPTRE). 
These standard methods should have false positives when not including covariates (the confounders) and should have none or few false positives when including the covariates.





# Load


```{r output=FALSE}
# Libraries and Parameters
# ========================================================================================================================================================================
suppressPackageStartupMessages(library(assertthat)) # for some assert statements
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(ggh4x))


theme_set(theme_cowplot() +
            theme(plot.title = element_text(hjust = .5),
                  plot.subtitle = element_text(hjust = .5),
                  strip.background = element_rect(fill = "white", color = 'black'),
                  panel.grid.minor = element_blank(),
                  panel.grid.major = element_blank()))


# === === === === === === === === === === === === === === === === === === === === === === === === === 
# parameters
DEVICE                = 'macbook'
AYZW_setting_name     = 'A'
NC_names_raw = 'PCA-singlegene'
NC_names = strsplit(x = NC_names_raw, split = '-') |> unlist()
proximal_setting_name_count = 'many'
NC_names_chosen = NC_names # just choose all here
source('../PATHS.R') # load in data_dir and save_dir and CODE_DIR, depending on DEVICE value

# --- Folder ---
# plot_dir = sprintf('%s/AY/%s/plots/basicProx/', save_dir, AYZW_setting_name)
# dir.create(plot_dir, showWarnings = FALSE, recursive=TRUE)

plot_dir = sprintf('%s/causarray/plots/', save_dir)
dir.create(plot_dir, showWarnings = FALSE)

# === === === === === === === === === === === === === === === === === === === === === === === === === 

# load chosen AYZW names
AY   = read.csv(sprintf('%s/AY/%s/AY.csv', save_dir, AYZW_setting_name))

# Continuous
# ========================================================================================================================================================================
# effects_continuous = read.csv(sprintf('%s/AY/%s/continuous-%s/effects_continuous.csv', save_dir, AYZW_setting_name, NC_names_raw))
# effects_continuous = effects_continuous |> filter(type != 'maybe')
# effects_continuous$type = factor(effects_continuous$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


# --- Resulting Data ---
chosen_NCtype_numNC = data.frame(NC_type=c('poisYA', 'poisYAU', 'nbYA', 'nbYAU', 'lmYA', 'lmYAU', 'PCA', 'SPCA8.0', 'SPCA34.5', 'WGCNA', 'singlegene'),
                                 numNC  =c(     NA ,        NA,    NA ,      NA,    NA ,      NA,   10,         15,         10,       9,      25))
chosen_NCtype_numNC = rbind(chosen_NCtype_numNC, 
                            data.frame(NC_type=c('proximalNegBinCountCount', 'proximalNegBinSinglegeneSinglegene', 'proximalNegBinPCAPCA'),
                                       numNC  =c(           10             ,                  10                 ,            10            ))
                           )


effects_continuous = read.csv(sprintf('%s/AY/%s/continuous-%s/effects_continuous.csv', save_dir, AYZW_setting_name, NC_names_raw))
# effects_continuous = effects_continuous |> filter(type != 'maybe')
# effects_continuous$type = factor(effects_continuous$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))

# --- Colors and Ordering and Format Text Labels ---


# # config colors for plot
# method_long_colors = c( colorRampPalette(c('white', '#00a300'))(8)[c(4, 8)], 
#                         viridis::plasma(11)[c(3, 4, 5, 6, 7, 8, 9, 10)])
# barplot(height = 1:length(method_long_colors), col = method_long_colors)
# for each NC_type, choose just 1 numNC to represent 
use_all_numNC = FALSE
chosen_numNC_subset = list('PCA'        = c(1, 3, 5, 10, 20),
                           'singlegene' = c(1, 3, 5, 10, 20))

color_pallette = paletteer::paletteer_d("colorblindr::OkabeIto_black") 
NC_base_colors = c('lm' = '#C11800FF', 'PCA'=color_pallette[3], 'SPCA8.0'=color_pallette[6], 'SPCA34.5'=color_pallette[1], 'WGCNA'=color_pallette[5], 'singlegene'=color_pallette[7])
# NC_base_colors = c('PCA'='blue', 'SPCA8.0'='red', 'SPCA34.5'='orange', 'WGCNA'='green', 'singlegene'='purple')
barplot(names.arg = NC_base_colors, height = 1:length(NC_base_colors), col = NC_base_colors)

format_methodlong <- function(method, NC_type, numNC) {
  if(method == 'lmYA' )     {return("Lin Reg")}
  if(method == 'lmYAU')     {return("Lin Reg (w/ confounders)")}
  if(method == 'poisYA')    {return("Pois")}
  if(method == 'poisYAU')   {return("Pois (w/ confounders)")}
  if(method == 'negbinYA')  {return("Neg Bin")}
  if(method == 'negbinYAU') {return("Neg Bin (w/ confounders)")}
  
  if(method == '2SLSpci2s') {
    if(       NC_type == 'proximalNegBinCountCount') {
      NC_type_short = 'Neg Bin Y w/ count gene NCs'
    } else if(NC_type == 'proximalNegBinSinglegeneSinglegene') {
      NC_type_short = 'Neg Bin Y w/ continuous gene NCs'
    } else if(NC_type == 'proximalNegBinPCAPCA') {
      NC_type_short = 'Neg Bin Y w/ continuous PCA NCs'
    } else {
      NC_type_short = NC_type
    }
    return(sprintf('Proximal: %s (#NC=%d)', NC_type_short, numNC))
  }
}


# source(sprintf('%s/AY/proximal_continuous_settings.r', save_dir)) # loads in list: proximal_continuous_settings

# construct the ordering
method_long_paste  =  c('lmYAlmYANA', 'lmYAUlmYAUNA')
method_long_order  =  c('Lin Reg', 'Lin Reg (w/ confounders)')
method_long_colors =  colorRampPalette(c('white', NC_base_colors[['lm']]))(8)[c(4, 8)]
for(NC_name in NC_names_chosen) {
  # proximal_setting = readRDS(sprintf('%s/AY/%s/%s/proximal_setting.rds', save_dir, AYZW_setting_name, NC_name))
  # proximal_setting = proximal_continuous_settings[[NC_name]]
  # cur_numNCs = proximal_setting$num_NC_pairs
  # if only select a few numNCs, select that here
  all_cur_numNCS =  effects_continuous |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
  if(use_all_numNC) {
    cur_numNCs = all_cur_numNCS
  } else {
    cur_numNCs = intersect(chosen_numNC_subset[[NC_name]], all_cur_numNCS)
  }
  
  
  colors_to_add = colorRampPalette(c('white', NC_base_colors[NC_name]))(length(cur_numNCs) + 1)[-(1)]
  method_long_names_to_add = c()
  for(numNC_1 in cur_numNCs) {
    method_long_names_to_add = c(method_long_names_to_add, format_methodlong('2SLSpci2s', NC_name, numNC_1))
  }
  
  method_long_paste  = c(method_long_paste, paste0('2SLSpci2s', NC_name,cur_numNCs))
  method_long_order  = c(method_long_order, method_long_names_to_add)
  method_long_colors = c(method_long_colors, colors_to_add)
  rm( method_long_names_to_add, colors_to_add)
}
names(method_long_colors) = method_long_order
methodlong_continuous_df = data.frame(level = method_long_paste,
                                      label = method_long_order, 
                                      color = method_long_colors)


effects_continuous = effects_continuous |> 
                     mutate(method_long = factor(paste0(method, NC_type, numNC), 
                                                 levels = method_long_paste,
                                                 labels = method_long_order))

effects_continuous = effects_continuous |> mutate(method = factor(method, levels = c('lmYA', 'lmYAU', '2SLSpci2s')))


#' convert NC_type to the current method_i, if NC_type is NA and method_i is lmYA or lmYAU
fn_turn_NC_type_lm <- function(NC_type_i, method_i) {
  if(is.na(NC_type_i) && method_i %in% c('lmYA', 'lmYAU')) {return(method_i)}
  else {return(NC_type_i)}
}
effects_continuous$NC_type = mapply(FUN = fn_turn_NC_type_lm,
                                    NC_type_i = effects_continuous$NC_type, 
                                    method_i = as.character(effects_continuous$method)) |> unname()
effects_continuous = effects_continuous |> mutate(NC_type = factor(NC_type, levels = c('lmYA', 'lmYAU', NC_names_chosen)))


dim(effects_continuous); head(effects_continuous)



# Count GLM
# ========================================================================================================================================================================
effects_count = read.csv(sprintf('%s/AY/%s/count-%s/effects_count.csv', save_dir, AYZW_setting_name, proximal_setting_name_count)) 
effects_count = effects_count #  |> filter(type != 'maybe')
# effects_count = effects_count |> filter(type != 'maybe')
# effects_count$type = factor(effects_count$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


method_long_level_count = c() # the original paste(..,..,..) for method long column
method_long_label_count = c() # the nice labels for method long column (also contains the order)
method_long_color_count =  c() # the colors for method long count methods

# Add count GLMs
method_long_level_count = c(method_long_level_count, c( "poisYANANA",  "poisYAUNANA", "nbYANANA", "nbYAUNANA"))
method_long_label_count = c(method_long_label_count, c("Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)"))
method_long_color_count = c(method_long_color_count, 
                            c(colorRampPalette(c('white', 'darkblue'))(8)[c(4, 8)],    #light blue, darker blue
                              colorRampPalette(c('white', 'darkorchid4'))(8)[c(4, 8)]  #light teal, darker teal
                            ))

chosen_numNC_subset_count = list(
                           'proximalNegBinCountCount'              = c(1, 3, 5, 10, 20),
                           'proximalNegBinSinglegeneSinglegene'    = c(1, 3, 5, 10, 20),
                           'proximalNegBinPCAPCA'                  = c(1, 3, 5, 10, 20))
color_pallette_count = paletteer::paletteer_d("colorblindr::OkabeIto_black") 
color_pallette_count = c('orange', 'pink', 'darkgreen') # temp colors
NC_base_colors_count = c('proximalNegBinCountCount'          =color_pallette_count[1],
                         'proximalNegBinSinglegeneSinglegene'=color_pallette_count[2], 
                         'proximalNegBinPCAPCA'              =color_pallette_count[3])
for(NC_name in names(chosen_numNC_subset_count)) {
  # print(paste0(length(method_long_level_count), '-', length(method_long_label_count)))
  all_cur_numNCS =  effects_count |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
  cur_numNCs = intersect(chosen_numNC_subset_count[[NC_name]], all_cur_numNCS)
  
  if(length(cur_numNCs) == 0) { # this method was excluded here
    next
  }
  colors_to_add = colorRampPalette(c('white', NC_base_colors_count[NC_name]))(length(cur_numNCs) + 1)[-(1)]
  method_long_names_to_add = c()
  for(numNC_1 in cur_numNCs) {
    method_long_names_to_add = c(method_long_names_to_add, format_methodlong('2SLSpci2s', NC_name, numNC_1))
  }
  
  method_long_level_count  = c(method_long_level_count, paste0('2SLSpci2s', NC_name, cur_numNCs))
  method_long_label_count  = c(method_long_label_count, method_long_names_to_add)
  method_long_color_count = c(method_long_color_count, colors_to_add)
  rm( method_long_names_to_add, colors_to_add, cur_numNCs, all_cur_numNCS, NC_name, numNC_1)
}

names(method_long_color_count) = method_long_label_count
# put into a dataframe to be cleaner
methodlong_count_df = data.frame(level = method_long_level_count,
                                 label = method_long_label_count, 
                                 color = method_long_color_count)

effects_count = effects_count |> 
                    mutate(method_long = factor(paste0(method, NC_type, numNC), 
                                                levels = methodlong_count_df$level, 
                                                labels = methodlong_count_df$label))


# effects_count |> filter(is.na(method_long))
# effects_continuous |> filter(is.na(method_long))

dim(effects_count); head(effects_count)


# SCEPTRE
# ========================================================================================================================================================================
# read in the previously formatted relevant rows when creating the AY pairs (2.3_xxx.r)
effects_sceptre = read.csv(sprintf('%s/AY/%s/sceptre_AY.csv', save_dir, AYZW_setting_name)) 
# effects_sceptre = effects_sceptre |> filter(type != 'maybe')
# effects_sceptre$type = factor(effects_sceptre$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


# head(effects_sceptre); colnames(effects_sceptre)
# colnames(effects_count)
# format to other effects df format
# doesn't really matter, but for now, we set the
# ATE estimate = log_2_fold_change! and the se is NA for now.
# because we have se(fold_change) but maybe needs some more transofrmation for log_2
effects_sceptre = effects_sceptre |>
  dplyr::rename(pval = p_value, ATE = log_2_fold_change) |>
  dplyr::mutate(se = NA, time_sec = NA, NC_type = NA,
                numNC = NA, basis = NA
                )

effects_sceptre = effects_sceptre[, colnames(effects_sceptre) %in% colnames(effects_continuous)]


method_long_colors_sceptre = c('SCEPTRE'='purple1', 'SCEPTRE (w/ confounders)'='purple4')
methodlong_colors_sceptre_df = data.frame(level = c('sceptreNANA', 'sceptreUNANA'),
                                           label = names(method_long_colors_sceptre),
                                           color = method_long_colors_sceptre,
                                           row.names = names(method_long_colors_sceptre))




# Config order and colors
# ========================================================================================================================================================================
# # for each method, color that represents. also use this ordering
# method_color_df = data.frame(method=c('lmYA', 'lmYAU', 
#                                         'poisYA', 'poisYAU', 
#                                         'nbYA', 'nbYAU', 
#                                         'sceptre', 'sceptreU'),
#                               color =c(colorRampPalette(c('white', 'darkblue'))(8)[c(4, 8)],   
#                                        colorRampPalette(c('white', 'darkgreen'))(8)[c(4, 8)],
#                                        colorRampPalette(c('white', 'darkorchid4'))(8)[c(4, 8)],
#                                        colorRampPalette(c('white', 'darkorange2'))(8)[c(4, 8)]
#                              ))
# 
# # 2 lines for (w/ confounders)
# method_nicenames_multiline = c('lmYA'    = 'Linear',
#                      'lmYAU'   = 'Linear\n(w/ confounders)',
#                      'poisYA'  = 'Poisson',
#                      'poisYAU' = 'Poisson\n(w/ confounders)',
#                      'nbYA'    = 'Negative Binomial',
#                      'nbYAU'   = 'Negative Binomial\n(w/ confounders)',
#                      'sceptre' = 'SCEPTRE',
#                      'sceptreU' = 'SCEPTRE\n(w/ confounders)')
# # 1 line
# method_nicenames = c('lmYA'    = 'Linear', 
#                      'lmYAU'   = 'Linear (w/ confounders)', 
#                      'poisYA'  = 'Poisson', 
#                      'poisYAU' = 'Poisson (w/ confounders)', 
#                      'nbYA'    = 'Negative Binomial', 
#                      'nbYAU'   = 'Negative Binomial (w/ confounders)', 
#                      'sceptre' = 'SCEPTRE',
#                      'sceptreU' = 'SCEPTRE (w/ confounders)')
# method_color_df$method_nicename = unname(method_nicenames)
# 
# method_colors = method_color_df$color |> `names<-`(method_color_df$method_nicename)


# config colors for plot
method_long_colors_all = c(method_long_colors[1:2], 
                           method_long_color_count[1:4], 
                           method_long_colors[3:length(method_long_colors)],
                           method_long_color_count[5:length(method_long_color_count)], 
                           method_long_colors_sceptre)
barplot(1:length(method_long_colors_all), col = method_long_colors_all)


methodlong_colors_all_df = rbind(methodlong_continuous_df,
                                  methodlong_count_df,
                                  methodlong_colors_sceptre_df)
# barplot(1:length(method_color_df$color), col = method_color_df$color)


# All
# ========================================================================================================================================================================
effects = dplyr::bind_rows(effects_continuous, effects_count, effects_sceptre)
# effects = effects |> filter(type != 'maybe')
effects$type = factor(effects$type,
       levels = c('negative',    'maybe',   'positive'),
       labels = c(    'Null', 'Discovery', 'Alternative'))

# effects = effects |> mutate(method_long = paste0(method, numNC))
# effects |> arrange(A, Y, method)

# effects = effects |> 
#           mutate(method_raw       = method, 
#                  method           = factor(method_raw, levels = method_color_df$method, labels = method_color_df$method_nicename),
#                  method_multiline = factor(method_raw, levels = names(method_nicenames_multiline), labels = method_nicenames_multiline)) 



effects = effects |> 
          mutate(method_long_raw = paste0(method, NC_type, numNC), 
                 method_long     = factor(method_long_raw, levels = methodlong_colors_all_df$level, labels = methodlong_colors_all_df$label)) 
effects = effects |> select(-basis)
```




```{r}
head(effects)
```

```{r}

get_AY_test_type <- function(isEqual) {
  if(isEqual) {
    return('positive')
  } else {
    return('discovery')
  }
}

# load causarray
# ========================================================================================================================================================================
# read in the previously formatted relevant rows when creating the AY pairs (2.3_xxx.r)
effects_causarray = read.csv(sprintf('%s/causarray/LFC.csv', save_dir)) 

effects_causarray = effects_causarray |>
  dplyr::rename(pval = pvalue, ATE = log2fc, se = log2fc_se, Y = gene_names, A=trt) |>
  dplyr::select(A, Y, ATE, se, pval) |> 
  dplyr::mutate(se = NA, time_sec = NA, NC_type = NA,
                numNC = NA, basis = NA
                )

grna_target_df = data.frame(A=unique(effects_causarray$A)) |> mutate(grna_target = sapply(A, FUN=function(x){substr(x, 1, nchar(x)-2)}))

effects_causarray = merge(grna_target_df, effects_causarray, by.x = 'A', by.y = 'A')


effects_causarray = effects_causarray |> mutate(type_bool = (grna_target==Y),
                                                type = sapply(type_bool, get_AY_test_type), .after=Y)


effects_causarray$type = factor(effects_causarray$type,
       levels = c('negative', 'discovery', 'positive'),
       labels = c(    'Null', 'Discovery', 'Alternative'))



effects_causarray = effects_causarray |> filter(!is.na(pval))

dim(effects_causarray)
```


# Plot causarray

Histogram of p-values

```{r fig.height=3, fig.width=6}

plot_df = effects_causarray

ggplot(plot_df) + 
  geom_histogram(aes(x = pval), color = 'black', fill = 'plum2', binwidth=.02) + 
  facet_wrap(vars(type), scales = 'free_y') +
  labs(title = 'Histogram of p-values from causarray')


ggsave(sprintf('%s/hist_pvals.pdf', plot_dir), height = 3, width= 6)
ggsave(sprintf('%s/hist_pvals.png', plot_dir), height = 3, width= 6, dpi=320)
```


Histogram of p-values of the tests that appear in other analyses 
```{r fig.height=3, fig.width=6}
plot_df = merge(effects |> select(A, Y) |> distinct(), effects_causarray, by = c('A', 'Y'))

ggplot(plot_df) + 
  geom_histogram(aes(x = pval), color = 'black', fill = 'plum2', binwidth=.02) + 
  facet_wrap(vars(type), scales = 'free_y') +
  labs(title = 'Histogram of p-values from causarray', 
       subtitle = 'filtered for AY tests that are also chosen in prox')

ggsave(sprintf('%s/hist_pvals_sameAYtests.pdf', plot_dir), height = 3, width= 6)
ggsave(sprintf('%s/hist_pvals_sameAYtests.png', plot_dir), height = 3, width= 6, dpi=320)
```

QQ-Plots (basically)


```{r fig.height=3, fig.width=6}
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects_causarray |> 
          group_by(type) |> 
          reframe(pval=sort(pval), count=sum(!is.na(pval)), rejection_rate = (1:count)/count) 

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .4, alpha = .9, key_glyph = "rect", color = 'plum2') +
  geom_point(size=.2, alpha=.7, key_glyph = "rect", color = 'plum2') +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  # scale_color_manual(values = method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate') +
  facet_grid(cols = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
# ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir), height = 3, width= 6)
ggsave(sprintf('%s/pval_qqplot.png', plot_dir), height = 3, width= 6, dpi = 600)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval+(1e-500), base = 10), x = -log(rejection_rate+(1e-500), base = 10))) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .4, alpha = .8, key_glyph = 'rect', color = 'plum2') +
  geom_point(size = 1, alpha = .7, key_glyph = 'rect', stroke=NA, color = 'plum2') +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  # scale_color_manual(values = method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)')) +
  facet_grid(cols = vars(type), scales = 'free') + # facetted_pos_scales doesn't seem to work with facet_grid
  # facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log.png', plot_dir), height = 3, width= 6, dpi = 600)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 2)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 2))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.png', plot_dir), height = 3, width= 6, dpi = 600)

#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + geom_line() + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 1)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 1)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 1))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 3), oob=scales::squish))
  )

ggsave(sprintf('%s/pval_qqplot_log_tail2.png', plot_dir), height = 3, width= 6, dpi = 600)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'


# ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir), height = 3, width= 9)
```


# Compare causarray to others


```{r fig.height=3, fig.width=9}

# chosen_method_long = c(
#       'Lin Reg',
#       'Lin Reg (w/ confounders)',
#       'Negative Binomial',
#       'Negative Binomial (w/ confounders)',
#       'SCEPTRE',
#       'SCEPTRE (w/ confounders)',
#       'Proximal: PCA (#NC=10)', 
#       'Proximal: singlegene (#NC=10)', 
#       'Proximal: Neg Bin Y w/ continuous PCA NCs (#NC=10)', 
#       'Proximal: Neg Bin Y w/ continuous gene NCs (#NC=10)'
# )

chosen_method_long_colors = c(
      # 'Lin Reg'='red1',
      'Lin Reg (w/ confounders)'='red4',
      # 'Negative Binomial'='purple1',
      'Negative Binomial (w/ confounders)'='purple4',
      # 'SCEPTRE'='blue1',
      'SCEPTRE (w/ confounders)'='blue4',
      'Causarray'='plum2',
      'Proximal: PCA (#NC=10)'='green1', 
      'Proximal: singlegene (#NC=10)'='green4', 
      'Proximal: Neg Bin Y w/ continuous PCA NCs (#NC=10)'='cyan1', 
      'Proximal: Neg Bin Y w/ continuous gene NCs (#NC=10)'='cyan4'
)


plot_df = dplyr::bind_rows(effects |> filter(type %in% c('Discovery', 'Alternative')) |>
                             filter(method_long %in% names(chosen_method_long_colors)),
                           effects_causarray |> mutate(method_long = 'Causarray')) |> 
  group_by(type, method_long) |> 
  reframe(pval=sort(pval), count=sum(!is.na(pval)), rejection_rate = (1:count)/count) |> 
  mutate(method_long = factor(method_long, levels = names(chosen_method_long_colors)))

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .4, alpha = .9, key_glyph = "rect") +
  geom_point(size=.2, alpha=.7, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = chosen_method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate', color = 'Method') +
  facet_grid(cols = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))

ggsave(sprintf('%s/compare_pval_qqplot.pdf', plot_dir), height = 3, width= 9)
ggsave(sprintf('%s/compare_pval_qqplot.png', plot_dir), height = 3, width= 9, dpi = 600)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval+(1e-500), base = 10), x = -log(rejection_rate+(1e-500), base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .4, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 1, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = chosen_method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  facet_grid(cols = vars(type), scales = 'free') + # facetted_pos_scales doesn't seem to work with facet_grid
  # facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p
ggsave(sprintf('%s/compare_pval_qqplot_log.pdf', plot_dir), height = 3, width= 9)
ggsave(sprintf('%s/compare_pval_qqplot_log.png', plot_dir), height = 3, width= 9, dpi = 600)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 2.5)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 2.5)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 2.5))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 9), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 9), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 9), oob=scales::squish))
  )
ggsave(sprintf('%s/compare_pval_qqplot_log_tail1.pdf', plot_dir), height = 3, width= 9)
ggsave(sprintf('%s/compare_pval_qqplot_log_tail1.png', plot_dir), height = 3, width= 9, dpi = 600)

#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + geom_line() + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(0, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(0, 2)),
    type == 'Discovery'   ~ scale_x_continuous(limits = c(0, 2))),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish),
    type == 'Discovery'   ~ scale_y_continuous(limits = c(0, 7), oob=scales::squish))
  )
ggsave(sprintf('%s/compare_pval_qqplot_log_tail2.pdf', plot_dir), height = 3, width= 9)
ggsave(sprintf('%s/compare_pval_qqplot_log_tail2.png', plot_dir), height = 3, width= 9, dpi = 600)




```







