---
title: "5.1_visualizeProxByNC"
format: 
  html:
    theme: flatly
    toc: true
    toc-depth: 4
    fig-format: svg 
    embed-resources: true
    self-contained-math: true
editor: source
execute: 
  echo: false
  warning: false
  error: false
---







# Load


```{r output=FALSE}
# Libraries and Parameters
# ========================================================================================================================================================================
suppressPackageStartupMessages(library(assertthat)) # for some assert statements
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(ggh4x))


theme_set(theme_cowplot() +
            theme(plot.title = element_text(hjust = .5),
                  plot.subtitle = element_text(hjust = .5),
                  strip.background = element_rect(fill = "white", color = 'black'),
                  panel.grid.minor = element_blank(),
                  panel.grid.major = element_blank()))


# === === === === === === === === === === === === === === === === === === === === === === === === === 
# parameters
DEVICE                = 'laptop'
AYZW_setting_name     = 'A'
NC_names_raw = 'PCA-SPCA8.0-SPCA34.5-WGCNA-singlegene'
NC_names = strsplit(x = NC_names_raw, split = '-') |> unlist()
NC_names_chosen = NC_names[c(1, 2, 3, 4, 5)] #all for now, but only include SPC8.0 later i think, can set the order of the NCs here

source('../PATHS.R') # load in data_dir and save_dir and CODE_DIR, depending on DEVICE value
# === === === === === === === === === === === === === === === === === === === === === === === === === 



# Continuous
# ========================================================================================================================================================================

# --- Folder ---
plot_dir_continuous = sprintf('%s/AY/%s/plots/continuous/', save_dir, AYZW_setting_name) # create plot_dir (based on save_dir + settings)
dir.create(plot_dir_continuous, showWarnings = FALSE, recursive=TRUE)

# --- Resulting Data ---
chosen_NCtype_numNC = data.frame(NC_type=c('poisYA', 'poisYAU', 'nbYA', 'nbYAU', 'lmYA', 'lmYAU', 'PCA', 'SPCA8.0', 'SPCA34.5', 'WGCNA', 'singlegene'),
                                 numNC  =c(     NA ,        NA,    NA ,      NA,    NA ,      NA,   10,         15,         10,       9,      25))


effects_continuous = read.csv(sprintf('%s/AY/%s/%s/effects_continuous.csv', save_dir, AYZW_setting_name, NC_names_raw))
effects_continuous = effects_continuous |> filter(type != 'maybe')
effects_continuous$type = factor(effects_continuous$type,
       levels = c('negative', 'positive'),
       labels = c(    'Null', 'Alternative'))

# --- Colors and Ordering and Format Text Labels ---


# # config colors for plot
# method_long_colors = c( colorRampPalette(c('white', '#00a300'))(8)[c(4, 8)], 
#                         viridis::plasma(11)[c(3, 4, 5, 6, 7, 8, 9, 10)])
# barplot(height = 1:length(method_long_colors), col = method_long_colors)
# for each NC_type, choose just 1 numNC to represent 
use_all_numNC = FALSE
chosen_numNC_subset = list('PCA'        = c(1, 3, 5,  7, 10, 15, 20),
                           'SPCA8.0'    = c(1, 3, 5,  7, 10, 15, 20),
                           'SPCA34.5'   = c(1, 3, 5,  7, 10, 15, 20),
                           'WGCNA'      = c(1, 3, 5,  7,  9),
                           'singlegene' = c(1, 3, 5, 10, 15, 30, 50))

color_pallette = paletteer::paletteer_d("colorblindr::OkabeIto_black") 
NC_base_colors = c('lm' = '#C11800FF', 'PCA'=color_pallette[3], 'SPCA8.0'=color_pallette[6], 'SPCA34.5'=color_pallette[1], 'WGCNA'=color_pallette[5], 'singlegene'=color_pallette[7])
# NC_base_colors = c('PCA'='blue', 'SPCA8.0'='red', 'SPCA34.5'='orange', 'WGCNA'='green', 'singlegene'='purple')
barplot(names.arg = NC_base_colors, height = 1:length(NC_base_colors), col = NC_base_colors)

format_methodlong <- function(method, NC_type, numNC) {
  if(method == 'lmYA' ) {return("Lin Reg")}
  if(method == 'lmYAU') {return("Lin Reg (w/ confounders)")}
  
  if(method == '2SLSpci2s') {
    return(sprintf('Proximal: %s (#NC=%d)', NC_type, numNC))
  }
}


# source(sprintf('%s/AY/proximal_continuous_settings.r', save_dir)) # loads in list: proximal_continuous_settings

# construct the ordering
method_long_paste  =  c('lmYANANA', 'lmYAUNANA')
method_long_order  =  c('Lin Reg', 'Lin Reg (w/ confounders)')
method_long_colors =  colorRampPalette(c('white', NC_base_colors[['lm']]))(8)[c(4, 8)]
for(NC_name in NC_names_chosen) {
  # proximal_setting = readRDS(sprintf('%s/AY/%s/%s/proximal_setting.rds', save_dir, AYZW_setting_name, NC_name))
  # proximal_setting = proximal_continuous_settings[[NC_name]]
  # cur_numNCs = proximal_setting$num_NC_pairs
  # if only select a few numNCs, select that here
  if(use_all_numNC) {
    cur_numNCs = effects_continuous |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
  } else {
    cur_numNCs = chosen_numNC_subset[[NC_name]]
  }
  
  
  colors_to_add = colorRampPalette(c('white', NC_base_colors[NC_name]))(length(cur_numNCs) + 1)[-(1)]
  method_long_names_to_add = c()
  for(numNC_1 in cur_numNCs) {
    method_long_names_to_add = c(method_long_names_to_add, format_methodlong('2SLSpci2s', NC_name, numNC_1))
  }
  
  method_long_paste  = c(method_long_paste, paste0('2SLSpci2s', NC_name,cur_numNCs))
  method_long_order  = c(method_long_order, method_long_names_to_add)
  method_long_colors = c(method_long_colors, colors_to_add)
  rm( method_long_names_to_add, colors_to_add)
}
names(method_long_colors) = method_long_order

effects_continuous = effects_continuous |> 
                     mutate(method_long = factor(paste0(method, NC_type, numNC), 
                                                 levels = method_long_paste,
                                                 labels = method_long_order))

effects_continuous = effects_continuous |> mutate(method = factor(method, levels = c('lmYA', 'lmYAU', '2SLSpci2s')))


#' convert NC_type to the current method_i, if NC_type is NA and method_i is lmYA or lmYAU
fn_turn_NC_type_lm <- function(NC_type_i, method_i) {
  if(is.na(NC_type_i) && method_i %in% c('lmYA', 'lmYAU')) {return(method_i)}
  else {return(NC_type_i)}
}
effects_continuous$NC_type = mapply(FUN = fn_turn_NC_type_lm,
                                    NC_type_i = effects_continuous$NC_type, 
                                    method_i = as.character(effects_continuous$method)) |> unname()
effects_continuous = effects_continuous |> mutate(NC_type = factor(NC_type, levels = c('lmYA', 'lmYAU', NC_names_chosen)))


dim(effects_continuous); head(effects_continuous)


# Count GLM
# ========================================================================================================================================================================
proximal_setting_name_count = 'simpleCount'


plot_dir_count = sprintf('%s/AY/%s/plots/count/', save_dir, AYZW_setting_name) # create plot_dir (based on save_dir + settings)
dir.create(plot_dir_count, showWarnings = FALSE)



# config colors for plot
method_long_colors_count = c(colorRampPalette(c('white', 'darkblue'))(8)[c(4, 8)], #light blue, darker blue
                             colorRampPalette(c('white', 'darkorchid4'))(8)[c(4, 8)]  #light teal, darker teal
                             )
barplot(1:length(method_long_colors_count), col = method_long_colors_count)
names(method_long_colors_count) =  c("Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)")

effects_count = read.csv(sprintf('%s/AY/%s/%s/effects_countGLM.csv', save_dir, AYZW_setting_name, proximal_setting_name_count)) # laptop avg time .08 vs ubergenno avg time .93
effects_count = effects_count |> mutate(method_long = paste0(method, numNC))

effects_count$method_long = factor(effects_count$method_long, 
       levels = c( "poisYANA",  "poisYAUNA", "nbYANA", "nbYAUNA"),
       labels = c("Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)"))



effects_count$type = factor(effects_count$type,
       levels = c('negative', 'positive'),
       labels = c(    'Null', 'Alternative'))


dim(effects_count); head(effects_count)


# All
# ========================================================================================================================================================================

# save results of 'all' together in continuous folder
plot_dir_all = sprintf('%s/AY/%s/plots/all/', save_dir, AYZW_setting_name)  # create plot_dir (based on save_dir + settings)
dir.create(plot_dir_all, showWarnings = FALSE)


# config colors for plot
method_long_colors_all = c(method_long_colors_count, method_long_colors)
barplot(1:length(method_long_colors_all), col = method_long_colors_all)

# bind df and order cols
NC_type_all_ordering = c('lmYA'    = 'Linear', 
                         'lmYAU'   = 'Linear\n(w/ confounders)', 
                         'poisYA'  = 'Poisson', 
                         'poisYAU' = 'Poisson\n(w/ confounders)', 
                         'nbYA'    = 'Negative Binomial', 
                         'nbYAU'   = 'Negative Binomial\n(w/ confounders)', 
                         NC_names_chosen |> `names<-`( NC_names_chosen ))


# effects = rbind(effects_continuous, effects_count)
effects = dplyr::bind_rows(effects_continuous, effects_count |> mutate(NC_type = method)) # and fill NA if col not present
effects$NC_type     = factor(effects$NC_type,     levels = names(NC_type_all_ordering))
effects$method_long = factor(effects$method_long, levels = names(method_long_colors_all))

# effects = effects |> mutate(method_long = paste0(method, numNC))
# 
# effects$method_long = factor(effects$method_long, 
#        levels = c("poisYANA",      "poisYAUNA", "nbYANA",           "nbYAUNA",                             "lmYANA",                            "lmYAUNA" ,       
#                          "2SLSpci2s1",       "2SLSpci2s3",       "2SLSpci2s5",       "2SLSpci2s8",       "2SLSpci2s10",       "2SLSpci2s15",       "2SLSpci2s20"),
#        labels = c( "Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)", "Lin Reg", "Lin Reg (w/ confounders)", 
#                    "Proximal (#NC=1)", "Proximal (#NC=3)", "Proximal (#NC=5)", "Proximal (#NC=8)", "Proximal (#NC=10)", "Proximal (#NC=15)", "Proximal (#NC=20)"))

# effects$type = factor(effects$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))

dim(effects); head(effects)

```


```{r}
print(sprintf('The NC_names_raw is: %s', NC_names_raw))
print(sprintf('The chosen NCs are: %s', paste0(NC_names, collapse = ', ')))
```



# Continuous Method Results

## Histogram of estimates and p-values

Histogram of estimates
```{r, fig.width=13, fig.height = 25}
ggplot(effects_continuous,
       aes(x = ATE, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .05) + # bins = 50) +
  geom_vline(aes(xintercept = 0), color = 'gray15', linewidth = .4) +
  coord_cartesian(xlim = c(-2, 1), ylim = c(0, 3)) +
  scale_y_continuous(breaks = 0:2, expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_long_colors, 
                    guide = guide_legend(ncol = 1)) +
  facet_grid(cols = vars(type), 
             rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC), 
             scales = 'free') + 
  labs(title = 'Histogram of Estimates',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))

ggsave(sprintf('%s/estimates_hist.pdf', plot_dir_continuous), height = 25, width= 13)
```


Histogram of p-values
```{r fig.width=13, fig.height = 25}

p = ggplot(effects_continuous,
       aes(x = pval, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .020) + # bins = 50) +
  scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1'), 
                     expand = expansion(mult = c(.01, .01))) +
  # coord_cartesian(ylim = c(0, 3)) +
  scale_y_continuous( expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_long_colors, guide = guide_legend(ncol = 1)) +
  facet_grid2(cols = vars(type), 
             rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC), 
             axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
             scales = 'free') + 
  labs(title = 'Histogram of p-values',
       x = 'p-value',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))
p

ggsave(sprintf('%s/pval_hist.pdf', plot_dir_continuous), height = 25, width= 13)


# # # force same axes- does not look good, but gives perspective
# p + facet_grid2(cols = vars(type),
#              rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC),
#              axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
#              scales = 'fixed')
```



## QQ-plots/Rejection Rates of p-values


### all methods with all NCs


QQ-plot (basically)
```{r fig.width=12, fig.height = 6}
# config df for plotting
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects_continuous |> group_by(type, method_long, NC_type, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) 

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_point() +
  geom_line(linewidth = .4, alpha = .9, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(rows = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir_continuous), height = 7, width= 8)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_line(linewidth = .9, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_long_colors, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
  facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log.pdf', plot_dir_continuous), height = 7, width= 8)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.4, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(x = list(
  type == 'Null'        ~ scale_x_continuous(limits = c(1, 3.25)),
  type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))
))
ggsave(sprintf('%s/pval_qqplot_log_tail2.pdf', plot_dir_continuous), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(1, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 1.25))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 35)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 60))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir_continuous), height = 6, width= 6)
```









### all methods with one choice of NC

```{r fig.width=6, fig.height = 6}



# config df for plotting
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects_continuous |> group_by(type, method_long, NC_type, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) 
plot_df = merge(chosen_NCtype_numNC, plot_df, by = c('NC_type', 'numNC'))

# make these colors the NC_base_colors (e.g. the color that would've bene for the max num NC)
method_long_colors_1numNC =  c(method_long_colors['Lin Reg'], method_long_colors['Lin Reg (w/ confounders)'],
                               sapply(X = names(method_long_colors)[-(1:2)], FUN = function(v){NC_base_colors[strsplit(v, split = ' ')[[1]][2]] |> unname()}))




#----- p-values: qqplot vs uniform = rejection rate by pval threshold (subset chosen 1 numNC per NC_type) ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_point() +
  geom_line(linewidth = .8, alpha = .9, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_long_colors_1numNC, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(rows = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot_1numNC.pdf', plot_dir_continuous), height = 6, width= 6)



#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .5, alpha = .5, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_long_colors_1numNC, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
  facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log_1numNC.pdf', plot_dir_continuous), height = 6, width= 6)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.4, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1_1numNC.pdf', plot_dir_continuous), height = 6, width= 6)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(x = list(
  type == 'Null'        ~ scale_x_continuous(limits = c(1, 3.25)),
  type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))
))
ggsave(sprintf('%s/pval_qqplot_log_tail2_1numNC.pdf', plot_dir_continuous), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(1, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 1.25))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 35)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 60))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail3_1numNC.pdf', plot_dir_continuous), height = 6, width= 6)

```


### each method with all choices of NC


#### qqplot (regular)


```{r fig.width=6, fig.height = 6}

make_qqplot <- function(plot_df) {
  #----- p-values: qqplot vs uniform = rejection rate by pval threshold ---------- 'pval_qqplot.pdf'
  ggplot(plot_df,
         aes(x = pval, y = rejection_rate, color = method_long)) +
    geom_abline(slope = 1, intercept = 0) +
    # geom_point() +
    geom_line(linewidth = 1, alpha = .95, key_glyph = "rect") +
    scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_color_manual(values = method_long_colors) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
    labs(title = 'Rejection Rate by P-value Threshold', 
         x = 'p-value threshold', y = 'rejection rate',
         color = 'Method') +
    facet_grid(rows = vars(type)) + 
    theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
}


# focus only on PCA
make_qqplot(effects_continuous |> filter(NC_type == 'PCA' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/PCA_pval_qqplot.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on SPCA8.0 
make_qqplot(effects_continuous |> filter(NC_type == 'SPCA8.0' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/SPCA8.0_pval_qqplot.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on SPCA34.5 
make_qqplot(effects_continuous |> filter(NC_type == 'SPCA34.5' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/SPCA34.5_pval_qqplot.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on WGCNA 
make_qqplot(effects_continuous |> filter(NC_type == 'WGCNA' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/WGCNA_pval_qqplot.pdf', plot_dir_continuous), height = 6, width= 6)

# focus only on singlegene 
make_qqplot(effects_continuous |> filter(NC_type == 'singlegene' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/singlegene_pval_qqplot.pdf', plot_dir_continuous), height = 6, width= 6)




```


#### qqplot (log)

```{r fig.width=12, fig.height = 6}


make_qqplot_log <- function(plot_df) {
    p = ggplot(plot_df,
               aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
          geom_abline(slope = 1, intercept = 0) +
          geom_line(linewidth = .5, alpha = .5, key_glyph = 'rect') +
          geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
          # scale_x_continuous(limits = c(2, 3.5)) +
          # scale_y_continuous(limits = c(0, 200)) +
          scale_color_manual(values = method_long_colors) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
          labs(title = '-log(p-value) by -log(Unif(0,1))', 
               x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
               color = 'Method') +
          # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
          facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
          theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
  
    return(list(p=p, 
              ptail=p + ggh4x::facetted_pos_scales(
                            x = list(
                              type == 'Null'        ~ scale_x_continuous(limits = c(.05, 2)),
                              type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 2))  ),
                            y = list(
                              type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
                              type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
                            )))
}


# focus only on PCA
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'PCA' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "PCA")


ggsave(plot = res$p,     filename = sprintf('%s/PCA_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/PCA_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)



# focus only on SPCA8.0 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'SPCA8.0' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "SPCA8.0")
ggsave(plot = res$p,     filename = sprintf('%s/SPCA8.0_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/SPCA8.0_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on SPCA34.5 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'SPCA34.5' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "SPCA34.5")
ggsave(plot = res$p,     filename = sprintf('%s/SPCA34.5_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/SPCA34.5_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on WGCNA 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'WGCNA' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "WGCNA")
ggsave(plot = res$p,     filename = sprintf('%s/WGCNA_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/WGCNA_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)

# focus only on singlegene 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'singlegene' | NC_type == 'lmYA' | NC_type == 'lmYAU') |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "singlegene")
ggsave(plot = res$p,     filename = sprintf('%s/singlegene_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/singlegene_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)




```







# Count GLM Method Results






## Histogram of Estimates and p-values



Histogram of estimates
```{r, fig.width=15, fig.height = 5}
# horizantal version (looks better)
ggplot(effects_count,
       aes(x = ATE, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)),  bins = 40) + # binwidth = .25) +
  geom_vline(aes(xintercept = 0), color = 'gray10', linewidth = .2) +
  scale_fill_manual(values = method_long_colors_count) +
  facet_grid(rows = vars(type), cols = vars(method_long), scales = 'fixed') + 
  labs(title = 'Histogram of Estimates',
       fill = 'Method') +
  theme(strip.text.x = element_text(vjust = 1), 
        panel.spacing.x = unit(.15, "lines"))

ggsave(sprintf('%s/estimates_hist_hori.pdf', plot_dir_count), height = 5, width= 15)


# vertical version- to be aligned ith continuous version (hich has to be vert bc of ho many methods there are)
p = ggplot(effects_count |> 
             mutate(method_long = factor(method, 
                                         levels = c( "poisYA",  "poisYAU", "nbYA", "nbYAU"),
                                         labels = c("Poisson", "Poisson\n(w/ confounders)", "Negative Binomial", "Negative Binomial\n(w/ confounders)"))),
       aes(x = ATE, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)),  bins = 40) + # binwidth = .25) +
  geom_vline(aes(xintercept = 0), color = 'gray10', linewidth = .2) +
  scale_fill_manual(values = method_long_colors_count) +
  facet_grid(rows = vars(type), cols = vars(method_long), scales = 'fixed') + 
  labs(title = 'Histogram of Estimates',
       fill = 'Method') +
  theme(strip.text.x = element_text(vjust = 1))


ggsave(plot = p + facet_grid(cols = vars(type), rows = vars(method_long), scales = 'fixed') +
                  theme(strip.text.x = element_text(vjust = .5), 
                        strip.text.y = element_text(vjust = 1)), 
       filename = sprintf('%s/estimates_hist_vert.pdf', plot_dir_count), height = 8, width= 8)
```



Histogram of p-values
```{r fig.width=15, fig.height = 5}
ggplot(effects_count,
       aes(x = pval, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .02) +
  scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_fill_manual(values = method_long_colors_count) +
  facet_grid(rows = vars(type), cols = vars(method_long), scales = 'fixed') +
  labs(title = 'Histogram of p-values',
       x = 'p-value',
       fill = 'Method') +
  theme(strip.text.x = element_text(vjust = 1))

ggsave(sprintf('%s/pval_hist_hori.pdf', plot_dir_count), height = 5, width= 15)

# vertical version- copy paste prev plot except change method_long col- add changes hen saving
p = ggplot(effects_count  |> 
             mutate(method_long = factor(method, 
                                         levels = c( "poisYA",  "poisYAU", "nbYA", "nbYAU"),
                                         labels = c("Poisson", "Poisson\n(w/ confounders)", "Negative Binomial", "Negative Binomial\n(w/ confounders)"))),
       aes(x = pval, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .02) +
  scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_fill_manual(values = method_long_colors_count) +
  facet_grid(rows = vars(type), cols = vars(method_long), scales = 'fixed') +
  labs(title = 'Histogram of p-values',
       x = 'p-value',
       fill = 'Method') +
  theme(strip.text.x = element_text(vjust = 1))

ggsave(plot = p + facet_grid(cols = vars(type), rows = vars(method_long), scales = 'fixed') +
                  theme(strip.text.x = element_text(vjust = .5), 
                        strip.text.y = element_text(vjust = 1)),
       filename = sprintf('%s/pval_hist_vert.pdf', plot_dir_count), height = 8, width= 8)



```


## QQ-Plots of p-values

QQ-plot (basically)
```{r}
# config df for plotting
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects_count |> mutate(method_long = paste0(method, numNC)) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n())
plot_df$method_long = factor(plot_df$method_long, 
       levels = c( "poisYANA",  "poisYAUNA", "nbYANA", "nbYAUNA"),
       labels = c("Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)"))

# plot_df$type = factor(plot_df$type,
#        levels = c('negative', 'positive'),
#        labels = c(    'Null', 'Alternative'))


#----- p-values: qqplot vs uniform = rejection rate by pval threshold ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_point() +
  geom_line(linewidth = 1, alpha = .9) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_long_colors_count) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(rows = vars(type)) 

ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir_count), height = 6, width= 8)


#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_line(linewidth = .9, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_long_colors_count) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
  facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right")
p

ggsave(sprintf('%s/pval_qqplot_log.pdf', plot_dir_count), height = 6, width= 8)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.4, 2.5)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2.5))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 300)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 300))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.pdf', plot_dir_count), height = 6, width= 8)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(x = list(
  type == 'Null'        ~ scale_x_continuous(limits = c(1, 3.25)),
  type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))
))
ggsave(sprintf('%s/pval_qqplot_log_tail2.pdf', plot_dir_count), height = 6, width= 8)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(1, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 1.25))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 35)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 60))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir_count), height = 6, width= 8)
```






# Continuous + Count Results together 






## Computation Time

Histogram of computation time
```{r, fig.height=4, fig.width=6}



p = ggplot(effects, 
       aes(x = time_sec, y = method_long, fill = method_long)) +
  geom_boxplot(outliers = FALSE, box.linewidth = .2, median.linewidth = .4, whisker.linewidth = .4) +
  scale_fill_manual(values = method_long_colors_all) +
  scale_y_discrete(limits = rev) +
  # facet_grid(cols = vars(type)) +
  labs(title = 'Computation Time by Method',
       fill = 'Method', x = 'time (sec)', y = 'Method') +
  theme(legend.position = 'none', axis.text.y = element_text(hjust = 1, size = 5.5))

p

ggsave(sprintf('%s/time_boxplot.pdf', plot_dir_all), height = 4, width = 6)

p + scale_x_log10(labels = scales::label_number(accuracy = 1, big.mark = ''))
ggsave(sprintf('%s/time_boxplot_log.pdf', plot_dir_all), height = 4, width = 6)



```

## Histogram of estimates and p-values

Histogram of estimates
```{r, fig.width=13, fig.height = 23}
ggplot(effects |> mutate(NC_type = factor(NC_type, labels = NC_type_all_ordering)),
       aes(x = ATE, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .05) + # bins = 50) +
  geom_vline(aes(xintercept = 0), color = 'gray15', linewidth = .4) +
  coord_cartesian(xlim = c(-2, 1), ylim = c(0, 3)) +
  scale_y_continuous(breaks = 0:2, expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_long_colors_all, 
                    guide = guide_legend(ncol = 1)) +
  facet_grid(cols = vars(type), 
             rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC), 
             scales = 'free') + 
  labs(title = 'Histogram of Estimates',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 7), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))


ggsave(sprintf('%s/estimates_hist.pdf', plot_dir_all),  height = 32, width= 13)
```





Histogram of p-values
```{r fig.width=13, fig.height = 23}
# ggplot(effects,
#        aes(x = pval, fill = method_long)) +
#   geom_histogram(aes(y = after_stat(..density..)), binwidth = .02) +
#   scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
#   scale_fill_manual(values = method_long_colors_all) +
#   facet_grid(rows = vars(type), cols =  vars(method, numNC), scales = 'free') +
#   labs(title = 'Histogram of p-values',
#        x = 'p-value',
#        fill = 'Method')
# 
# ggsave(sprintf('%s/pval_hist.pdf', plot_dir_all), height = 5, width= 20)


p = ggplot(effects  |> mutate(NC_type = factor(NC_type, labels = NC_type_all_ordering)),
       aes(x = pval, fill = method_long)) +
  geom_histogram(aes(y = after_stat(..density..)), binwidth = .020) + # bins = 50) +
  scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1'), 
                     expand = expansion(mult = c(.01, .01))) +
  # coord_cartesian(ylim = c(0, 3)) +
  scale_y_continuous( expand = expansion(mult = c(0, .05))) +
  scale_fill_manual(values = method_long_colors_all, guide = guide_legend(ncol = 1)) +
  facet_grid2(cols = vars(type), 
             rows = vars(NC_type, numNC), # vars(method_long), # vars(method, NC_type, numNC), 
             axes = 'x', remove_labels = 'x', independent = 'y', shrink = TRUE,
             scales = 'free') + 
  labs(title = 'Histogram of p-values',
       x = 'p-value',
       fill = 'Method') +
  theme(strip.text.y = element_text(size = 7), 
        axis.text.y = element_text(size = 8), 
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(0, "lines"))
p

ggsave(sprintf('%s/pval_hist.pdf', plot_dir_all), height = 32, width= 13)

```

## QQ-plots/Rejection Rates of p-values

### all methods with all NCs


QQ-plot (basically)
```{r fig.width=6, fig.height = 6}
# config df for plotting
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects |> group_by(type, method_long, NC_type, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) 

#----- p-values: qqplot vs uniform = rejection rate by pval threshold (all numNC)  ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_point() +
  geom_line(linewidth = .4, alpha = .9, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_long_colors_all, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(rows = vars(type)) +
  theme(legend.text = element_text(size = 7), legend.key.size = unit(.3, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)




#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_line(linewidth = .9, alpha = .8, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_long_colors_all, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
  facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 7), legend.key.size = unit(.3, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log.pdf', plot_dir_all), height = 6, width= 6)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) ------------------------ 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.4, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1.pdf', plot_dir_all), height = 6, width= 6)


#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(x = list(
  type == 'Null'        ~ scale_x_continuous(limits = c(1, 3.25)),
  type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))
))
ggsave(sprintf('%s/pval_qqplot_log_tail2.pdf', plot_dir_all), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail3.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(1, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 1.25))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 35)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 60))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail3.pdf', plot_dir_all), height = 6, width= 6)
```









### all methods with one choice of NC

```{r fig.width=6, fig.height = 6}



# config df for plotting
# the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n)
plot_df = effects |> group_by(type, method_long, NC_type, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) 
plot_df = merge(chosen_NCtype_numNC, plot_df, by = c('NC_type', 'numNC'))

# make these colors the NC_base_colors (e.g. the color that would've bene for the max num NC)
method_long_colors_1numNC =  c(method_long_colors_all['Poisson'], method_long_colors_all['Poisson (w/ confounders)'],
                               method_long_colors_all['Negative Binomial'], method_long_colors_all['Negative Binomial (w/ confounders)'],
                               method_long_colors_all['Lin Reg'], method_long_colors_all['Lin Reg (w/ confounders)'],
                               sapply(X = names(method_long_colors)[-(1:2)], FUN = function(v){NC_base_colors[strsplit(v, split = ' ')[[1]][2]] |> unname()}))




#----- p-values: qqplot vs uniform = rejection rate by pval threshold (subset chosen 1 numNC per NC_type) ---------- 'pval_qqplot.pdf'
ggplot(plot_df,
       aes(x = pval, y = rejection_rate, color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  # geom_point() +
  geom_line(linewidth = .8, alpha = .8, key_glyph = "rect") +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = method_long_colors_1numNC, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = 'Rejection Rate by P-value Threshold', 
       x = 'p-value threshold', y = 'rejection rate',
       color = 'Method') +
  facet_grid(rows = vars(type)) +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
ggsave(sprintf('%s/pval_qqplot_1numNC.pdf', plot_dir_all), height = 6, width= 6)



#----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf'
p = ggplot(plot_df,
       aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_line(linewidth = .5, alpha = .5, key_glyph = 'rect') +
  geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
  # scale_x_continuous(limits = c(2, 3.5)) +
  # scale_y_continuous(limits = c(0, 200)) +
  scale_color_manual(values = method_long_colors_1numNC, guide = guide_legend(ncol = 1)) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
  labs(title = '-log(p-value) by -log(Unif(0,1))', 
       x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
       color = 'Method') +
  # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
  facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
  theme(legend.text = element_text(size = 8), legend.key.size = unit(.4, units = 'cm'))
p

ggsave(sprintf('%s/pval_qqplot_log_1numNC.pdf', plot_dir_all), height = 6, width= 6)



# use gg4x to set limits separately 
# https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975
#----- p-values: log10 qqplot (Focus on same tail) --------------------------------------- 'pval_qqplot_log_tail1.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.05, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 2))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail1_1numNC.pdf', plot_dir_all), height = 6, width= 6)


#----- p-values: log10 qqplot (more zoomed in) ------------------------------------------- 'pval_qqplot_log_tail2.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.05, 2)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 2))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 10)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 10))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail2_1numNC.pdf', plot_dir_all), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail3.pdf'
p + ggh4x::facetted_pos_scales(x = list(
  type == 'Null'        ~ scale_x_continuous(limits = c(1, 3.25)),
  type == 'Alternative' ~ scale_x_continuous(limits = c(.4, 2))
))
ggsave(sprintf('%s/pval_qqplot_log_tail3_1numNC.pdf', plot_dir_all), height = 6, width= 6)

#----- p-values: log10 qqplot (Focus even more on different tails) ------------------------ 'pval_qqplot_log_tail4.pdf'
p + ggh4x::facetted_pos_scales(
  x = list(
    type == 'Null'        ~ scale_x_continuous(limits = c(.5, 2.5)),
    type == 'Alternative' ~ scale_x_continuous(limits = c(.2, 1.7))  ),
  y = list(
    type == 'Null'        ~ scale_y_continuous(limits = c(0, 60)),
    type == 'Alternative' ~ scale_y_continuous(limits = c(0, 60))  )
  )

ggsave(sprintf('%s/pval_qqplot_log_tail4_1numNC.pdf', plot_dir_all), height = 6, width= 6)



```








### each method with all choices of NC


#### qqplot (regular)


```{r fig.width=6, fig.height = 6}

make_qqplot <- function(plot_df) {
  #----- p-values: qqplot vs uniform = rejection rate by pval threshold ---------- 'pval_qqplot.pdf'
  ggplot(plot_df,
         aes(x = pval, y = rejection_rate, color = method_long)) +
    geom_abline(slope = 1, intercept = 0) +
    # geom_point() +
    geom_line(linewidth = 1, alpha = .95, key_glyph = "rect") +
    scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_color_manual(values = method_long_colors_all) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
    labs(title = 'Rejection Rate by P-value Threshold', 
         x = 'p-value threshold', y = 'rejection rate',
         color = 'Method') +
    facet_grid(rows = vars(type)) + 
    theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
}




# focus only on PCA
make_qqplot(effects |> filter(NC_type == 'PCA' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/PCA_pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)


# focus only on SPCA8.0 
make_qqplot(effects |> filter(NC_type == 'SPCA8.0' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/SPCA8.0_pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)


# focus only on SPCA34.5 
make_qqplot(effects |> filter(NC_type == 'SPCA34.5' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/SPCA34.5_pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)


# focus only on WGCNA 
make_qqplot(effects |> filter(NC_type == 'WGCNA' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/WGCNA_pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)

# focus only on singlegene 
make_qqplot(effects |> filter(NC_type == 'singlegene' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
ggsave(sprintf('%s/singlegene_pval_qqplot.pdf', plot_dir_all), height = 6, width= 6)




```


#### qqplot (log)

```{r fig.width=12, fig.height = 6}


make_qqplot_log <- function(plot_df) {
    p = ggplot(plot_df,
               aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
          geom_abline(slope = 1, intercept = 0) +
          geom_line(linewidth = .5, alpha = .5, key_glyph = 'rect') +
          geom_point(size = 2, alpha = .7, key_glyph = 'rect', stroke=NA) +
          # scale_x_continuous(limits = c(2, 3.5)) +
          # scale_y_continuous(limits = c(0, 200)) +
          scale_color_manual(values = method_long_colors_all) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
          labs(title = '-log(p-value) by -log(Unif(0,1))', 
               x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
               color = 'Method') +
          # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
          facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
          theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
  
    return(list(p=p, 
              ptail=p + ggh4x::facetted_pos_scales(
                            x = list(
                              type == 'Null'        ~ scale_x_continuous(limits = c(.05, 2)),
                              type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 2))  ),
                            y = list(
                              type == 'Null'        ~ scale_y_continuous(limits = c(0, 40)),
                              type == 'Alternative' ~ scale_y_continuous(limits = c(0, 40))  )
                            )))
}


# focus only on PCA
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'PCA' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "PCA")
ggsave(plot = res$p,     filename = sprintf('%s/PCA_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/PCA_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)



# focus only on SPCA8.0 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'SPCA8.0' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "SPCA8.0")
ggsave(plot = res$p,     filename = sprintf('%s/SPCA8.0_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/SPCA8.0_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on SPCA34.5 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'SPCA34.5' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "SPCA34.5")
ggsave(plot = res$p,     filename = sprintf('%s/SPCA34.5_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/SPCA34.5_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)


# focus only on WGCNA 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'WGCNA' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "WGCNA")
ggsave(plot = res$p,     filename = sprintf('%s/WGCNA_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/WGCNA_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)

# focus only on singlegene 
res = make_qqplot_log(effects_continuous |> filter(NC_type == 'singlegene' | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) )
gridExtra::grid.arrange(res$p, res$ptail, nrow = 1, top = "singlegene")
ggsave(plot = res$p,     filename = sprintf('%s/singlegene_pval_qqplot_log.pdf',       plot_dir_continuous), height = 6, width= 6)
ggsave(plot = res$ptail, filename = sprintf('%s/singlegene_pval_qqplot_log_tail1.pdf', plot_dir_continuous), height = 6, width= 6)




```
















# Focus on 1 NC type

## Focus on PCA

Edit the plots to showcase the results. This will be specific for each setting. For example, we can crop to highlight the tails and subset for different methods

```{r fig.height=6, fig.width=12}
NC_name = 'PCA'

cur_numNCs =  effects_continuous |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
# combine original count GLM cols. redo NC_name cols to be have better scale (e.g. viridis vs white to blue)
method_long_colors_best = c(method_long_colors_all[1:6], # keep count GLM and lin reg colors
                            viridis::viridis(2 + length(cur_numNCs))[2 + 1:length(cur_numNCs)] |> rev() |> `names<-` ( grep(pattern=' PCA ', x=names(method_long_colors_all), value=TRUE)  ) )
# scales::show_col(method_long_colors_best)

# focus only on PCA
plot_df = effects |> filter(NC_type == NC_name | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n())




p = ggplot(plot_df,
               aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
          geom_abline(slope = 1, intercept = 0) +
          geom_line(linewidth = .9, alpha = .85, key_glyph = 'rect') +
          # geom_point(size = 1.2, alpha = .7, key_glyph = 'rect', stroke=NA) +
          # scale_x_continuous(limits = c(2, 3.5)) +
          # scale_y_continuous(limits = c(0, 200)) +
          scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
          labs(title = '-log(p-value) by -log(Unif(0,1))', 
               x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
               color = 'Method') +
          # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
          facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
          theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
ptail = p + ggh4x::facetted_pos_scales(
                            x = list(
                              type == 'Null'        ~ scale_x_continuous(limits = c(.05, 1.75)),
                              type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 1.75))  ),
                            y = list(
                              type == 'Null'        ~ scale_y_continuous(limits = c(0, 20)),
                              type == 'Alternative' ~ scale_y_continuous(limits = c(0, 20))  )
                            )  
ggsave(plot = p,     filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log.pdf',       plot_dir_all, NC_name), height = 6, width= 6)
ggsave(plot = ptail, filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log_tail1.pdf', plot_dir_all, NC_name), height = 6, width= 6)

gridExtra::grid.arrange(p, ptail, nrow = 1, top = NC_name)
```





## Focus on SPCA8.0



```{r fig.height=6, fig.width=12}
NC_name = 'SPCA8.0'

cur_numNCs =  effects_continuous |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
# combine original count GLM cols. redo NC_name cols to be have better scale (e.g. viridis vs white to blue)
method_long_colors_best = c(method_long_colors_all[1:6], # keep count GLM and lin reg colors
                            viridis::viridis(2 + length(cur_numNCs))[2 + 1:length(cur_numNCs)] |> rev() |> `names<-` ( grep(pattern=NC_name, x=names(method_long_colors_all), value=TRUE)  ) )
# scales::show_col(method_long_colors_best)

# focus only on NC_name
plot_df = effects |> filter(NC_type == NC_name | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n())




p = ggplot(plot_df,
               aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
          geom_abline(slope = 1, intercept = 0) +
          geom_line(linewidth = .9, alpha = .85, key_glyph = 'rect') +
          # geom_point(size = 1.2, alpha = .7, key_glyph = 'rect', stroke=NA) +
          # scale_x_continuous(limits = c(2, 3.5)) +
          # scale_y_continuous(limits = c(0, 200)) +
          scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
          labs(title = '-log(p-value) by -log(Unif(0,1))', 
               x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
               color = 'Method') +
          # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
          facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
          theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
ptail = p + ggh4x::facetted_pos_scales(
                            x = list(
                              type == 'Null'        ~ scale_x_continuous(limits = c(.05, 1.75)),
                              type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 1.75))  ),
                            y = list(
                              type == 'Null'        ~ scale_y_continuous(limits = c(0, 20)),
                              type == 'Alternative' ~ scale_y_continuous(limits = c(0, 20))  )
                            )  
ggsave(plot = p,     filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log.pdf',       plot_dir_all, NC_name), height = 6, width= 6)
ggsave(plot = ptail, filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log_tail1.pdf', plot_dir_all, NC_name), height = 6, width= 6)

gridExtra::grid.arrange(p, ptail, nrow = 1, top = NC_name)
```



## Focus on singlegenes



```{r fig.height=6, fig.width=12}
NC_name = 'singlegene'

cur_numNCs =  effects_continuous |> filter(NC_type == NC_name) |> pull(numNC) |> unique() |> sort()
# combine original count GLM cols. redo NC_name cols to be have better scale (e.g. viridis vs white to blue)
method_long_colors_best = c(method_long_colors_all[1:6], # keep count GLM and lin reg colors
                            viridis::viridis(2 + length(cur_numNCs))[2 + 1:length(cur_numNCs)] |> rev() |> `names<-` ( grep(pattern=NC_name, x=names(method_long_colors_all), value=TRUE)  ) )
# scales::show_col(method_long_colors_best)

# focus only on NC_name
plot_df = effects |> filter(NC_type == NC_name | NC_type %in% c('lmYA', 'lmYAU', 'poisYA', 'poisYAU', 'nbYA', 'nbYAU')) |> group_by(type, method_long) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n())




p = ggplot(plot_df,
               aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) +
          geom_abline(slope = 1, intercept = 0) +
          geom_line(linewidth = .9, alpha = .85, key_glyph = 'rect') +
          # geom_point(size = 1.2, alpha = .7, key_glyph = 'rect', stroke=NA) +
          # scale_x_continuous(limits = c(2, 3.5)) +
          # scale_y_continuous(limits = c(0, 200)) +
          scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) +
          labs(title = '-log(p-value) by -log(Unif(0,1))', 
               x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'),
               color = 'Method') +
          # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid
          facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") +
          theme(legend.text = element_text(size = 9), legend.key.size = unit(.4, units = 'cm'))
ptail = p + ggh4x::facetted_pos_scales(
                            x = list(
                              type == 'Null'        ~ scale_x_continuous(limits = c(.05, 1.75)),
                              type == 'Alternative' ~ scale_x_continuous(limits = c(.05, 1.75))  ),
                            y = list(
                              type == 'Null'        ~ scale_y_continuous(limits = c(0, 20)),
                              type == 'Alternative' ~ scale_y_continuous(limits = c(0, 20))  )
                            )  
ggsave(plot = p,     filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log.pdf',       plot_dir_all, NC_name), height = 6, width= 6)
ggsave(plot = ptail, filename = sprintf('%s/FOCUS_ON_%s_pval_qqplot_log_tail1.pdf', plot_dir_all, NC_name), height = 6, width= 6)

gridExtra::grid.arrange(p, ptail, nrow = 1, top = NC_name)
```



<!-- ## Best Plots (Focusing on SPCA 8.0) -->


<!-- Edit the plots to showcase the results. This will be specific for each setting. For example, we can crop to highlight the tails and subset for different methods (e.g. not all NCs because that can be too many and hard to see). -->


<!-- ```{r} -->
<!-- if(proximal_setting_name_continuous == 'SPCA8.0') { -->
<!--   # redo the colors bc we subset for numNC so we can choose better -->
<!--   method_long_colors_best = c(method_long_colors_count, # keep count GLM and lin reg colors -->
<!--                               method_long_colors[c(1, 2)], -->
<!--                               viridis::plasma(8)[c(3, 4, 5, 6)]) -->

<!--   barplot(1:length(method_long_colors_best), col = method_long_colors_best) -->
<!-- } -->

<!-- ``` -->



<!-- Histogram of Estimates -->
<!-- ```{r fig.width=15, fig.height = 5} -->
<!-- if(proximal_setting_name_continuous == 'SPCA8.0') { -->
<!--   p = ggplot(effects_continuous |> filter(is.na(numNC) | numNC %in% c(5, 10, 15, 20)), -->
<!--          aes(x = ATE, fill = method_long)) + -->
<!--     geom_histogram(aes(y = after_stat(..density..)), binwidth = .15) + -->
<!--     geom_vline(aes(xintercept = 0), color = 'gray20', linewidth = .4) + -->
<!--     scale_x_continuous(limits = c(-2.25, 2.25)) + -->
<!--     scale_fill_manual(values = method_long_colors_best[-(1:4)]) + -->
<!--     facet_grid(rows = vars(type), cols = vars(method_long), scales = 'free') + -->
<!--     labs(title = 'Histogram of Estimates', -->
<!--          fill = 'Method') -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/estimates_hist_2.pdf', plot_dir_all), height = 5, width= 16) -->
<!-- } -->
<!-- ``` -->


<!-- Histogram of p-values -->
<!-- ```{r fig.width=15, fig.height = 5} -->
<!-- if(proximal_setting_name_continuous == 'SPCA8.0') { -->
<!--   p = ggplot(effects |> filter(is.na(numNC) | numNC %in% c(5, 10, 15, 20)) |> -->
<!--            mutate(method_long_facet = factor(paste0(method, numNC), -->
<!--          levels = c("poisYANA",      "poisYAUNA", "nbYANA",           "nbYAUNA",                             "lmYANA",                            "lmYAUNA" , -->
<!--                            "2SLSpci2s1",       "2SLSpci2s3",       "2SLSpci2s5",       "2SLSpci2s8",      "2SLSpci2s10",       "2SLSpci2s15",       "2SLSpci2s20"), -->
<!--          labels = c( "Poisson\n", "Poisson\n(w/ confounders)", "Negative Binomial\n", "Negative Binomial\n(w/ confounders)", "Linear Reg\n", "Linear Reg\n(w/ confounders)", -->
<!--                      "Proximal\n(#NC=1)", "Proximal\n(#NC=3)", "Proximal\n(#NC=5)", "Proximal\n(#NC=8)", "Proximal\n(#NC=10)", "Proximal\n(#NC=15)", "Proximal\n(#NC=20)"))), -->
<!--          aes(x = pval, fill = method_long)) + -->
<!--     geom_histogram(aes(y = after_stat(..density..)), binwidth = .05) + -->
<!--     scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) + -->
<!--     scale_fill_manual(values = method_long_colors_best) + -->
<!--     facet_grid(rows = vars(type), cols = vars(method_long_facet), scales = 'free') + -->
<!--     labs(title = 'Histogram of p-values', -->
<!--          x = 'p-value', -->
<!--          fill = 'Method') -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/pval_hist_2.pdf', plot_dir_all), height = 5, width= 25) -->
<!-- } -->
<!-- ``` -->







<!-- QQ-plot (basically) -->
<!-- ```{r fig.height = 6, fig.width= 8} -->
<!-- if(proximal_setting_name_continuous == 'SPCA8.0') { -->
<!--   # config df for plotting -->
<!--   # the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n) -->
<!--   plot_df = effects |> mutate(method_long = paste0(method, numNC)) |> group_by(type, method_long, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) -->
<!--   plot_df$method_long = factor(plot_df$method_long, -->
<!--          levels = c("poisYANA",      "poisYAUNA", "nbYANA",           "nbYAUNA",                             "lmYANA",                            "lmYAUNA" , -->
<!--                            "2SLSpci2s1",       "2SLSpci2s3",       "2SLSpci2s5",       "2SLSpci2s8",      "2SLSpci2s10",       "2SLSpci2s15",       "2SLSpci2s20"), -->
<!--          labels = c( "Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)", "Lin Reg", "Lin Reg (w/ confounders)", -->
<!--                      "Proximal (#NC=1)", "Proximal (#NC=3)", "Proximal (#NC=5)", "Proximal (#NC=8)", "Proximal (#NC=10)", "Proximal (#NC=15)", "Proximal (#NC=20)")) -->

<!--   plot_df = plot_df |> filter(is.na(numNC) | numNC %in% c(5, 10, 15, 20)) -->
<!--   # plot_df$type = factor(plot_df$type, -->
<!--   #        levels = c('negative', 'positive'), -->
<!--   #        labels = c(    'Null', 'Alternative')) -->


<!--   #----- p-values: qqplot vs uniform = rejection rate by pval threshold ---------- 'pval_qqplot.pdf' -->
<!--   p = ggplot(plot_df, -->
<!--          aes(x = pval, y = rejection_rate, color = method_long)) + -->
<!--     geom_abline(slope = 1, intercept = 0) + -->
<!--     # geom_point() + -->
<!--     geom_line(linewidth = 1, alpha = .8, key_glyph = 'rect') + -->
<!--     scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) + -->
<!--     scale_y_continuous(limits = c(0, 1)) + -->
<!--     scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) + -->
<!--     labs(title = 'Rejection Rate by P-value Threshold', -->
<!--          x = 'p-value threshold', y = 'rejection rate', -->
<!--          color = 'Method') + -->
<!--     facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/pval_qqplot_2.pdf', plot_dir_all), height = 7.5, width = 7.5) -->


<!--   #----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf' -->
<!--   p = ggplot(plot_df, -->
<!--          aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) + -->
<!--     geom_abline(slope = 1, intercept = 0) + -->
<!--     # geom_line(linewidth = .3, alpha = .8) + -->
<!--     geom_point(size = 1.5, alpha = .8, key_glyph = 'rect') + -->
<!--     # scale_x_continuous(limits = c(2, 3.5)) + -->
<!--     # scale_y_continuous(limits = c(0, 200)) + -->
<!--     scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) + -->
<!--     labs(title = '-log(p-value) by -log(Unif(0,1))', -->
<!--          x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'), -->
<!--          color = 'Method') + -->
<!--     # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid -->
<!--     facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") -->


<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/pval_qqplot_log_best.pdf', plot_dir_all), height = 7.5, width = 7.5) -->

<!--   # use gg4x to set limits separately -->
<!--   # https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975 -->
<!--   #----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail_best.pdf' -->
<!--   p2 = p + ggh4x::facetted_pos_scales( -->
<!--     x = list( -->
<!--       type == 'Null'        ~ scale_x_continuous(limits = c(.35, 1.5)), -->
<!--       type == 'Alternative' ~ scale_x_continuous(limits = c(.35, 1.5))  ), -->
<!--     y = list( -->
<!--       type == 'Null'        ~ scale_y_continuous(limits = c(0, 25)), -->
<!--       type == 'Alternative' ~ scale_y_continuous(limits = c(0, 25))   ) -->
<!--     ) -->
<!--   print(p2) -->
<!--   ggsave(plot = p2, filename =sprintf('%s/pval_qqplot_log_tail_best.pdf', plot_dir_all), height = 7.5, width = 7.5) -->

<!-- } -->

<!-- ``` -->




<!-- ## Best Plots (Focusing on PCA) -->


<!-- Edit the plots to showcase the results. This will be specific for each setting. For example, we can crop to highlight the tails and subset for different methods (e.g. not all NCs because that can be too many and hard to see). -->


<!-- ```{r} -->
<!-- if(proximal_setting_name_continuous == 'PCA') { -->
<!--   # redo the colors bc we subset for numNC so we can choose better -->
<!--   method_long_colors_best = c(method_long_colors_count, # keep count GLM and lin reg colors -->
<!--                               method_long_colors[c(1, 2)], -->
<!--                               viridis::plasma(8)[c(3, 4, 5, 6)]) -->

<!--   barplot(1:length(method_long_colors_best), col = method_long_colors_best) -->
<!-- } -->

<!-- ``` -->



<!-- Histogram of Estimates -->
<!-- ```{r fig.width=15, fig.height = 5} -->
<!-- if(proximal_setting_name_continuous == 'PCA') { -->
<!--   p = ggplot(effects_continuous |> filter(is.na(numNC) | numNC %in% c(3, 5, 10, 20)), -->
<!--          aes(x = ATE, fill = method_long)) + -->
<!--     geom_histogram(aes(y = after_stat(..density..)), binwidth = .15) + -->
<!--     geom_vline(aes(xintercept = 0), color = 'gray20', linewidth = .4) + -->
<!--     scale_x_continuous(limits = c(-2.25, 2.25)) + -->
<!--     scale_fill_manual(values = method_long_colors_best[-(1:4)]) + -->
<!--     facet_grid(rows = vars(type), cols = vars(method_long), scales = 'free') + -->
<!--     labs(title = 'Histogram of Estimates', -->
<!--          fill = 'Method') -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/estimates_hist_2.pdf', plot_dir_all), height = 5, width= 16) -->
<!-- } -->
<!-- ``` -->


<!-- Histogram of p-values -->
<!-- ```{r fig.width=15, fig.height = 5} -->
<!-- if(proximal_setting_name_continuous == 'PCA') { -->
<!--   p = ggplot(effects |> filter(is.na(numNC) | numNC %in% c(3, 5, 10, 20)) |> -->
<!--            mutate(method_long_facet = factor(paste0(method, numNC), -->
<!--          levels = c("poisYANA",      "poisYAUNA", "nbYANA",           "nbYAUNA",                             "lmYANA",                            "lmYAUNA" , -->
<!--                            "2SLSpci2s1",       "2SLSpci2s3",       "2SLSpci2s5",       "2SLSpci2s8",      "2SLSpci2s10",       "2SLSpci2s15",       "2SLSpci2s20"), -->
<!--          labels = c( "Poisson\n", "Poisson\n(w/ confounders)", "Negative Binomial\n", "Negative Binomial\n(w/ confounders)", "Linear Reg\n", "Linear Reg\n(w/ confounders)", -->
<!--                      "Proximal\n(#NC=1)", "Proximal\n(#NC=3)", "Proximal\n(#NC=5)", "Proximal\n(#NC=8)", "Proximal\n(#NC=10)", "Proximal\n(#NC=15)", "Proximal\n(#NC=20)"))), -->
<!--          aes(x = pval, fill = method_long)) + -->
<!--     geom_histogram(aes(y = after_stat(..density..)), binwidth = .05) + -->
<!--     scale_x_continuous(limits = c(-.05, 1.05), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) + -->
<!--     scale_fill_manual(values = method_long_colors_best) + -->
<!--     facet_grid(rows = vars(type), cols = vars(method_long_facet), scales = 'free') + -->
<!--     labs(title = 'Histogram of p-values', -->
<!--          x = 'p-value', -->
<!--          fill = 'Method') -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/pval_hist_2.pdf', plot_dir_all), height = 5, width= 25) -->
<!-- } -->
<!-- ``` -->







<!-- QQ-plot (basically) -->
<!-- ```{r fig.height = 6, fig.width= 8} -->
<!-- if(proximal_setting_name_continuous == 'PCA') { -->
<!--   # config df for plotting -->
<!--   # the 'rejection_rate' column is also the 'theoretical' for the uniform (1/n, 2/n, ..., n/n) -->
<!--   plot_df = effects |> mutate(method_long = paste0(method, numNC)) |> group_by(type, method_long, numNC) |> reframe(pval=sort(pval), rejection_rate = (1:n())/n()) -->
<!--   plot_df$method_long = factor(plot_df$method_long, -->
<!--          levels = c("poisYANA",      "poisYAUNA", "nbYANA",           "nbYAUNA",                             "lmYANA",                            "lmYAUNA" , -->
<!--                            "2SLSpci2s1",       "2SLSpci2s3",       "2SLSpci2s5",       "2SLSpci2s8",      "2SLSpci2s10",       "2SLSpci2s15",       "2SLSpci2s20"), -->
<!--          labels = c( "Poisson", "Poisson (w/ confounders)", "Negative Binomial", "Negative Binomial (w/ confounders)", "Lin Reg", "Lin Reg (w/ confounders)", -->
<!--                      "Proximal (#NC=1)", "Proximal (#NC=3)", "Proximal (#NC=5)", "Proximal (#NC=8)", "Proximal (#NC=10)", "Proximal (#NC=15)", "Proximal (#NC=20)")) -->

<!--   plot_df = plot_df |> filter(is.na(numNC) | numNC %in% c(3, 5, 10, 20)) -->
<!--   # plot_df$type = factor(plot_df$type, -->
<!--   #        levels = c('negative', 'positive'), -->
<!--   #        labels = c(    'Null', 'Alternative')) -->


<!--   #----- p-values: qqplot vs uniform = rejection rate by pval threshold ---------- 'pval_qqplot.pdf' -->
<!--   p = ggplot(plot_df, -->
<!--          aes(x = pval, y = rejection_rate, color = method_long)) + -->
<!--     geom_abline(slope = 1, intercept = 0) + -->
<!--     # geom_point() + -->
<!--     geom_line(linewidth = 1, alpha = .8, key_glyph = 'rect') + -->
<!--     scale_x_continuous(limits = c(0, 1), breaks = c(0, .25, .5, .75, 1), labels = c('0', '.25', '.5', '.75', '1')) + -->
<!--     scale_y_continuous(limits = c(0, 1)) + -->
<!--     scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) + -->
<!--     labs(title = 'Rejection Rate by P-value Threshold', -->
<!--          x = 'p-value threshold', y = 'rejection rate', -->
<!--          color = 'Method') + -->
<!--     facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") -->
<!--   print(p) -->
<!--   ggsave(plot = p, filename =sprintf('%s/pval_qqplot_2.pdf', plot_dir_all), height = 7.5, width = 7.5) -->


<!--   #----- p-values: log10 qqplot -------------------------------------------------- 'pval_qqplot_log.pdf' -->
<!--   p = ggplot(plot_df, -->
<!--          aes(y = -log(pval, base = 10), x = -log(rejection_rate, base = 10), color = method_long)) + -->
<!--     geom_abline(slope = 1, intercept = 0) + -->
<!--     # geom_line(linewidth = .3, alpha = .8) + -->
<!--     geom_point(size = 1.5, alpha = .8, key_glyph = 'rect') + -->
<!--     # scale_x_continuous(limits = c(2, 3.5)) + -->
<!--     # scale_y_continuous(limits = c(0, 200)) + -->
<!--     scale_color_manual(values = method_long_colors_best) + #, labels = lapply(levels(plot_df$method_long), latex2exp::TeX)) + -->
<!--     labs(title = '-log(p-value) by -log(Unif(0,1))', -->
<!--          x = latex2exp::TeX('theoretical -$log_{10}$(Unif(0,1))'), y = latex2exp::TeX('sample -$log_{10}$(p-value)'), -->
<!--          color = 'Method') + -->
<!--     # facet_grid(rows = vars(type), scales = 'free') # facetted_pos_scales doesn't seem to work with facet_grid -->
<!--     facet_wrap(vars(type), ncol=1, scales = 'free', strip.position = "right") -->


<!--   print(p) -->
<!--   ggsave(plot = p, filename = sprintf('%s/pval_qqplot_log_best.pdf', plot_dir_all), height = 7.5, width = 7.5) -->

<!--   # use gg4x to set limits separately -->
<!--   # https://stackoverflow.com/questions/72962172/custom-y-axis-breaks-of-facet-wrap/73006975#73006975 -->
<!--   #----- p-values: log10 qqplot (Focus on different tails) ---------------------------------- 'pval_qqplot_log_tail_best.pdf' -->
<!--   p2 = p + ggh4x::facetted_pos_scales( -->
<!--     x = list( -->
<!--       type == 'Null'        ~ scale_x_continuous(limits = c(.35, 1.5)), -->
<!--       type == 'Alternative' ~ scale_x_continuous(limits = c(.35, 1.5))  ), -->
<!--     y = list( -->
<!--       type == 'Null'        ~ scale_y_continuous(limits = c(0, 25)), -->
<!--       type == 'Alternative' ~ scale_y_continuous(limits = c(0, 25))   ) -->
<!--     ) -->
<!--   print(p2) -->
<!--   ggsave(plot = p2, filename = sprintf('%s/pval_qqplot_log_tail_best.pdf', plot_dir_all), height = 7.5, width = 7.5) -->

<!-- } -->

<!-- ``` -->









